---
title: "Cell lines immunopeptidome analysis - Figure 3 - Supplementary file 1"
author: "Léa ROGUE"
output:
  pdf_document:
    fig_caption: true
    latex_engine: xelatex
    toc: true
    toc_depth: '4'
    includes:
      before_body: tex_files/list_of_figures.tex
date: "2026-02-02"
---

```{r, echo = FALSE, message=FALSE, warning=FALSE}
library(knitr)
opts_chunk$set(
  tidy.opts = list(width.cutoff = 60), 
  tidy = TRUE, 
  fig.align = 'center'
)
```

\newpage

# Load libraries and functions
```{r, results='hide', message=FALSE, warning=FALSE}
library(grid)
library(ggplot2)
library(reshape2)
library(UniProt.ws)
library(dplyr)
library(tidyr)
library(VennDiagram)
library(plotly)
library(ggrepel)
library(writexl)
```

```{r}
map_genes_to_peptides <- function(df_summary) {
  df_sep <- df_summary %>%
    mutate(RowID = row_number()) %>%
    separate_rows(Prot_accessions, sep = ";\\s*") %>%
    mutate(Prot_accessions = sub("-\\d+$", "", Prot_accessions))

  gene_mapping <- AnnotationDbi::select(
    up,
    keys    = unique(df_sep$Prot_accessions),
    columns = "gene_primary",
    keytype = "UniProtKB"
  )

  df_merged <- df_sep %>%
    left_join(gene_mapping, by = c("Prot_accessions" = "Entry"))

  df_final <- df_merged %>%
    group_by(Seq) %>%
    summarise(
      Proteins = paste(unique(Prot_accessions), collapse = "; "),
      Genes    = paste(unique(Gene.Names..primary.), collapse = "; "),
      .groups = "drop"
    )

  df_summary %>%
    left_join(df_final, by = "Seq") %>%
    filter(Genes %in% l_CTA)
}
prep_df <- function(df){

  df_summary <- df %>%
    group_by(Prot_accessions) %>%
    summarise(nb_sequences = n(), .groups = "drop")
  
  # Separate
  df_sep <- df_summary %>%
    mutate(RowID = row_number()) %>%
    separate_rows(Prot_accessions, sep = ";\\s*") %>%
    mutate(Prot_accessions = sub("-\\d+$", "", Prot_accessions))
  
  # Mapping UniProt -> Gene(s)
  gene_mapping <- AnnotationDbi::select(
    up,
    keys    = df_sep$Prot_accessions,
    columns = "gene_primary",
    keytype = "UniProtKB"
  )
  
  # Merge and reconstruct
  df_final <- df_sep %>%
    left_join(gene_mapping,
              by = c("Prot_accessions" = "Entry")) %>%
    group_by(RowID, nb_sequences) %>%
    summarise(
      Proteins = paste(unique(Prot_accessions), collapse = "; "),
      Genes    = paste(unique(Gene.Names..primary.), collapse = "; "),
      .groups = "drop"
    )%>%
    arrange(desc(nb_sequences))
  
  df_final <- df_final %>%
    group_by(Genes) %>%
    mutate(Genes_unique = if(n() > 1) paste0(Genes, "_", row_number()) else Genes) %>%
    ungroup() %>%
    mutate(Genes_unique = factor(Genes_unique, levels = Genes_unique)) %>%
    arrange(desc(nb_sequences))
  return(df_final)}
```

# Read data
```{r}
# Cell lines immunopeptidomic
df_ch_mhc1 <- read.table("../data/immunopep_CH2879.tsv", sep = "\t", header = T)
df_jj_mhc1 <- read.table("../data/immunopep_JJ012.tsv", sep = "\t", header = T)

# Filter on length
df_jj_mhc1 <- df_jj_mhc1[df_jj_mhc1$seq_len >= 8 & df_jj_mhc1$seq_len <= 11,]
df_ch_mhc1 <- df_ch_mhc1[df_ch_mhc1$seq_len >= 8 & df_ch_mhc1$seq_len <= 11,]

# CTA
l_CTA <- read.table("../data/CTA_list_clean.txt")$V1

# Read IEDB file
df_iedb_mhc1 <- read.csv("../data/epitope_table_export_human_MHC_I.tsv", header = T, sep = "\t")
df_pep_mhc1 <- df_iedb_mhc1[, c("Name", "Source.Molecule.IRI", "Molecule.Parent.IRI")]

# IEDB
pep_iedb_mhc1 <- unique(df_pep_mhc1$Name)
pep_iedb_mhc1 <- gsub(" .*|\\+.*", "", pep_iedb_mhc1)

# Read scatter data from Fig1E
df_scatter_hr <- read.table("../results/table_scatter_63_complete.tsv", sep = "\t", header = T)

# Read and prepare strong binders files
df_aff_jj <- read.table("../data/table_res_netmhc_all_JJ012.tsv", sep = "\t", header = T)
df_aff_ch <- read.table("../data/table_res_netmhc_all_CH2879.tsv", sep = "\t", header = T)
```


# I. Venn diagram with CTAs
```{r}
# CTA iedb
df_iedb_mhc1$ID <- gsub("http://www.uniprot.org/uniprot/", "", df_iedb_mhc1$Molecule.Parent.IRI)
df_iedb_mhc1$ID <- sub("-.*", "", df_iedb_mhc1$ID)

# Transform ID in gene name
up <- UniProt.ws(taxId = 9606)
gene_mapping <- AnnotationDbi::select(
  up,
  keys = unique(df_iedb_mhc1$ID[df_iedb_mhc1$ID != ""]) ,
  columns = "gene_primary",
  keytype = "UniProtKB"
)
colnames(gene_mapping) <- c("From", "ID", "Gene.Names..primary.")
df_iedb_mhc1_name <- merge(df_iedb_mhc1, gene_mapping, by.x = "ID", all.x = TRUE)

# Select peptides from CTAs of IEDB
df_iedb_mhc1_cta <- df_iedb_mhc1_name[df_iedb_mhc1_name$Gene.Names..primary. %in% l_CTA, ]
```


```{r}
# Take CTAs from peptides in cell lines
df_final_pep_jj <- map_genes_to_peptides(df_jj_mhc1)
df_final_pep_ch <- map_genes_to_peptides(df_ch_mhc1)

# Clean pep iedb
pep_iedb_mhc1_cta <- unique(df_iedb_mhc1_cta$Name)
pep_iedb_mhc1_cta <- gsub(" .*|\\+.*", "", pep_iedb_mhc1_cta)
pep_iedb_mhc1_cta <- unique(pep_iedb_mhc1_cta)

# Select new pep CH
new_pep_ch_mhc1 <- setdiff(df_final_pep_ch$Seq, pep_iedb_mhc1_cta)
grep_ch <- sapply(new_pep_ch_mhc1, function(x) any(grepl(x, pep_iedb_mhc1_cta)))
new_pep_ch_mhc1 <- new_pep_ch_mhc1[!new_pep_ch_mhc1 %in% names(grep_ch)[grep_ch]]
df_ch_new_pep_mhc1 <- df_ch_mhc1 %>%
  filter(Seq %in% new_pep_ch_mhc1) %>%
  distinct(Seq, .keep_all = TRUE)

# Select new pep JJ
new_pep_jj_mhc1 <- setdiff(df_final_pep_jj$Seq, pep_iedb_mhc1_cta)
grep_jj <- sapply(new_pep_jj_mhc1, function(x) any(grepl(x, pep_iedb_mhc1_cta)))
new_pep_jj_mhc1 <- new_pep_jj_mhc1[!new_pep_jj_mhc1 %in% names(grep_jj)[grep_jj]]

df_jj_new_pep_mhc1 <- df_jj_mhc1 %>%
  filter(Seq %in% new_pep_jj_mhc1) %>%
  distinct(Seq, .keep_all = TRUE)

# Clean IEDB pep
pep_iedb_extended_ch <- unique(c(pep_iedb_mhc1_cta, names(grep_ch)[grep_ch]))
pep_iedb_extended_jj <- unique(c(pep_iedb_mhc1_cta, names(grep_jj)[grep_jj]))
```



## JJ
```{r, fig.cap="Venn plot JJ012 (MHCI)"}
# Venn diagram
grid.newpage()
venn <- draw.pairwise.venn(
  area1 = length(pep_iedb_extended_jj),
  area2 = length(df_final_pep_jj$Seq),
  cross.area = length(df_final_pep_jj$Seq) - length(new_pep_jj_mhc1), 
  category = c("IEDB", "JJ012"),
  fill = c("#2C7BB6", "#ff7f0e"),
  scaled = FALSE)
grid.draw(venn)
```
\newpage

## CH
```{r, fig.cap="Venn plot CH2879 (MHCI)"}
grid.newpage()
venn <- draw.pairwise.venn(
  area1 = length(pep_iedb_extended_ch),
  area2 = length(df_final_pep_ch$Seq),
  cross.area = length(df_final_pep_ch$Seq) - length(new_pep_ch_mhc1), 
  category = c("IEDB", "CH2879"),
  fill = c("#2C7BB6", "#ff7f0e"),
  scaled = FALSE)
grid.draw(venn)
```
\newpage

## Take CTA peptides for further analysis
```{r}
# IEDB
df_iedb_cta <- df_iedb_mhc1_cta %>%
  mutate(Peptide = gsub(" .*|\\+.*", "", Name)) %>%
  filter(Peptide %in% pep_iedb_extended_ch | Peptide %in% pep_iedb_extended_jj) %>%
  dplyr::select(Gene = Gene.Names..primary., Peptide) %>%
  distinct(Gene, Peptide)

# JJ012
df_jj_cta <- df_final_pep_jj %>%
  filter(Seq %in% df_final_pep_jj$Seq) %>%
  dplyr::select(Gene = Genes, Peptide = Seq) %>%
  distinct(Gene, Peptide)

# CH2879
df_ch_cta <- df_final_pep_ch %>%
  dplyr::select(Gene = Genes, Peptide = Seq) %>%
  distinct(Gene, Peptide)

# Merge
df_cta_all <- bind_rows(df_iedb_cta, df_jj_cta, df_ch_cta) %>%
  distinct(Gene, Peptide) %>%
  arrange(Gene, Peptide)

fasta_lines <- c(rbind(paste0(">", seq_len(nrow(df_cta_all))), df_cta_all$Peptide))
writeLines(fasta_lines, "../results/peptides_CTA_uniques_jj_ch_iedb.fa")
write.table(df_cta_all, "../results/peptides_CTA_uniques_jj_ch_iedb.tsv", row.names = FALSE, sep = "\t", quote = FALSE)
```




# II. Histogram new peptides
## CH2879
```{r, fig.cap="CTA new peptides histogram (CH2879, MHCI)"}
# Prepare data
df_ch_new_pep_mhc1_plot <- prep_df(df_ch_new_pep_mhc1)

# Plot
df <- df_ch_new_pep_mhc1_plot[df_ch_new_pep_mhc1_plot$Genes %in% l_CTA,]
p <- ggplot(df, aes(x = Genes_unique, y = nb_sequences)) +
  geom_bar(stat = "identity", fill = "#ff7f0e") +
  theme_minimal() +
  labs(x = "Novel CTA proteins detected in CH2879",
       y = "Number of novel peptides") +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1, size = 5, color = "black"),
    panel.grid.major = element_line(color = "grey80", linewidth = 0.1),
    panel.grid.minor = element_line(color = "grey90", linewidth = 0.1)
  )+
  scale_x_discrete(labels = df$Genes) 
p
#ggsave("../results/histo_cta_ch.svg", plot = p, device = "svg", width = 4, height = 2)

```

\newpage


## JJ012
```{r, fig.cap="CTA new peptides histogram (JJ, MHCI)"}
# Prepare data
df_jj_new_pep_mhc1_plot <- prep_df(df_jj_new_pep_mhc1)

# Plot
df <- df_jj_new_pep_mhc1_plot[df_jj_new_pep_mhc1_plot$Genes %in% l_CTA,]
p <- ggplot(df, aes(x = Genes_unique, y = nb_sequences)) +
  geom_bar(stat = "identity", fill = "#ff7f0e") +
  theme_minimal() +
  labs(x = "Novel CTA proteins detected in JJ012",
       y = "Number of novel peptides") +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1, size = 5, color = "black"),
    panel.grid.major = element_line(color = "grey80", linewidth = 0.1),
    panel.grid.minor = element_line(color = "grey90", linewidth = 0.1)
  ) +
  scale_x_discrete(labels = df$Genes) 
p
#ggsave("../results/histo_cta_jj.svg", plot = p, device = "svg", width = 4.5, height = 2)
```
\newpage

# III. Peptide ranks and scatter plots

```{r}
# Select dominant HLA type
df_aff_ch_hlaa0101 <- df_aff_ch[df_aff_ch$Allele == "HLA-A0101" & df_aff_ch$Peptide %in% df_ch_mhc1$Seq, c("Peptide", "Affinity", "Rank")]

df_aff_jj_hlaa0201 <- df_aff_jj[df_aff_jj$Allele == "HLA-A0201" & df_aff_jj$Peptide %in% df_jj_mhc1$Seq, c("Peptide", "Affinity", "Rank")]

```

## CH2879
### Scatter plot from Fig1E
```{r, fig.cap="Scatter plot from Fig1E (CH2879)"}
# Add known pep
pep_commons <- intersect(df_final_pep_ch$Seq, pep_iedb_mhc1_cta)

# Reunite known peptides
pep_commons2 <- names(grep_ch)[grep_ch]
pep_known_all <- union(pep_commons, pep_commons2)

# Dataframe with know CTA peptides and new peptides
df_iedb_new_pep_ch <- df_ch_mhc1 %>%
  filter(Seq %in% new_pep_ch_mhc1 | Seq %in% pep_known_all) %>%
  distinct(Seq, .keep_all = TRUE)

# Prepare df for scatter
df_iedb_new_pep_ch_anno <- prep_df(df_iedb_new_pep_ch)
df_merge <- merge(df_iedb_new_pep_ch, df_iedb_new_pep_ch_anno, by.x = "Prot_accessions", by.y = "Proteins")

# Add rank information
df_pep_ch_hlaa0101 <- df_merge[df_merge$Seq %in% df_aff_ch_hlaa0101$Peptide,]
df_pep_ch_hlaa0101 <- merge(df_pep_ch_hlaa0101, df_aff_ch_hlaa0101, by.x = "Seq", by.y = "Peptide")
df_scatter_ch <- merge(df_scatter_hr, df_pep_ch_hlaa0101, by.x = "SYMBOL", by.y = "Genes")

# Plot
df_scatter_ch <- df_scatter_ch %>%
  mutate(Status = case_when(
    Seq %in% new_pep_ch_mhc1 ~ "New",
    Seq %in% pep_known_all   ~ "Known"))

# Agregate per gene, if one is new => orange
df_scatter_ch_gene <- df_scatter_ch %>%
  group_by(SYMBOL, Mean_expression_tissues, Intensity, HR) %>%
  summarise(
    Status = ifelse(any(Status == "New"), "New", "Known"),
    Rank = min(Rank),
    .groups = "drop"
  )

# Labeling genes
df_labels <- df_scatter_ch_gene %>%
  filter(Mean_expression_tissues < quantile(df_scatter_hr$Mean_expression_tissues, 0.75),
         Intensity > median(df_scatter_hr$Intensity)) %>%
  mutate(is_sb = Rank <= 0.5)


# For legends
df_legend <- data.frame(
  Mean_expression_tissues = c(NA, NA),
  Intensity = c(NA, NA),
  Status = c("Strong binders", "Others"))

# Plot
p <- ggplot(df_scatter_ch_gene,
            aes(x = Mean_expression_tissues,
                y = Intensity,
                color = Status,
                text = paste0("Gene: ", SYMBOL,
                              "<br>Status: ", Status,
                              "<br>Mean expression: ", round(Mean_expression_tissues, 3),
                              "<br>Intensity: ", round(Intensity, 3),
                              "<br>HR: ", round(HR, 3),
                              "<br>Rank: ", round(Rank, 3)))) +
  geom_point(size = 2, alpha = 0.8) +
  geom_point(
    data = df_legend,
    aes(x = Mean_expression_tissues, y = Intensity, color = Status),
    size = 0, alpha = 0, inherit.aes = FALSE) +
  scale_color_manual(
  values = c("Known" = "#1f77b4",
             "New" = "#ff7f0e",
             "Strong binders" = "black",
             "Others" = "black"),
  labels = c("Known" = "Known peptides",
             "New"  = "Novel peptides",
             "Strong binders" = expression(bold("Strong binders (bold)")),
             "Others" = "Others"),
  breaks = c("Known", "New", "Strong binders", "Others")) +
  guides(color = guide_legend(
    override.aes = list(
      size  = c(2, 2, 0, 0),
      alpha = c(1, 1, 0, 0)))) +
  scale_x_log10(breaks = c(0.001, 0.01, 0.1, 1, 10), labels = c("0.001", "0.01", "0.1", "1", "10")) +
  geom_hline(yintercept = median(df_scatter_hr$Intensity), linetype = "dashed", color = "black") +
  annotate("text", x = max(df_scatter_ch_gene$Mean_expression_tissues), y = median(df_scatter_hr$Intensity) - 0.2, label = "Median\nexpression", size = 2.5) +
  geom_vline(xintercept = quantile(df_scatter_hr$Mean_expression_tissues, 0.75), linetype = "dashed", color = "black") +
  annotate("text", x = quantile(df_scatter_hr$Mean_expression_tissues, 0.75) + 0.8, y = max(df_scatter_ch_gene$Intensity) + 0.2, label = "75th percentile", size = 2.5) +
  geom_text_repel(
    data = df_labels,
    aes(label = SYMBOL,
        fontface = ifelse(is_sb, "bold", "plain"),
        size     = 3),
    color         = "black",
    max.overlaps  = 20,
    box.padding   = 0.4,
    segment.color = "black",
    segment.size  = 0.3,
    min.segment.length = 0) +
  scale_size_identity() +
  theme_minimal() +
  labs(x = "CTAs expression in normal tissues (Log10 TPM mean)",
       y = "CTAs expression in CHS (mean intensities)",
       color = "CH2879 HLA-A*01:01\npeptides")
p
#ggplotly(p, tooltip = "text")
#ggsave("../results/scatter_plot_ch.svg", plot = p,  width = 20, height = 15, units = "cm")
```
\newpage

### Rank scatter plot
```{r, fig.cap="Rank scatter plot (CH2879)"}
# Count number of peptides per rank
df_count <- df_scatter_ch %>%
  ungroup() %>%
  mutate(Status = as.character(Status)) %>%
  group_by(Rank) %>%
  summarise(
    nb_pep = n(),
    is_new = any(Status == "New"))

# Scatter plot
p <- ggplot(df_count, aes(x = nb_pep, y = Rank, color = is_new)) +
  geom_point(size = 2) +
  scale_color_manual(values = c("FALSE" = "grey30", "TRUE" = "#ff7f0e"), 
                     labels = c("FALSE" = "Known peptides", "TRUE" = "Novel peptides")) +
  geom_hline(yintercept = c(0.5, 2), linetype = "dashed", color = "black") +
  annotate("text", x = max(df_count$nb_pep), y = 0.5 * 0.8, label = "Strong binders", hjust = 1, color = "black") +
  annotate("text", x = max(df_count$nb_pep), y = 2 * 0.8, label = "Weak binders", hjust = 1, color = "black") +
  labs(x = "CH2879 HLA-A*01:01 peptides number", y = "%Rank", color = "Category") +
  scale_y_log10(breaks = c(0.01, 0.1, 1, 10, 100), labels = c("0.01", "0.1", "1", "10", "100")) +
  scale_x_continuous(breaks = seq(0, max(df_count$nb_pep), by = 2)) +
  theme_minimal()
p
#ggsave("../results/scatter_ch_a01.svg", plot = p)
```
\newpage


## JJ012
### Scatter plot from Fig1E
```{r, fig.cap="Scatter plot from Fig1E (JJ012)"}
# Add known pep
pep_commons <- intersect(df_final_pep_jj$Seq, pep_iedb_mhc1_cta)

# Reunite known peptides
pep_commons2 <- names(grep_jj)[grep_jj]
pep_known_all <- union(pep_commons, pep_commons2)

# Dataframe with know CTA peptides and new peptides
df_iedb_new_pep_jj <- df_jj_mhc1 %>%
  filter(Seq %in% new_pep_jj_mhc1 | Seq %in% pep_known_all) %>%
  distinct(Seq, .keep_all = TRUE)

# Prepare df for scatter
df_iedb_new_pep_jj_anno <- prep_df(df_iedb_new_pep_jj)
df_merge <- merge(df_iedb_new_pep_jj, df_iedb_new_pep_jj_anno, by.x = "Prot_accessions", by.y = "Proteins")


# Add rank information
df_pep_jj_hlaa0201 <- df_merge[df_merge$Seq %in% df_aff_jj_hlaa0201$Peptide,]
df_pep_jj_hlaa0201 <- merge(df_pep_jj_hlaa0201, df_aff_jj_hlaa0201, by.x = "Seq", by.y = "Peptide")

# Prepare data for scatter
df_scatter_jj <- merge(df_scatter_hr, df_pep_jj_hlaa0201, by.x = "SYMBOL", by.y = "Genes")

# Plot
df_scatter_jj <- df_scatter_jj %>%
  mutate(Status = case_when(
    Seq %in% new_pep_jj_mhc1 ~ "New",
    Seq %in% pep_known_all   ~ "Known"))

# Agregate per gene, if one is new => orange
df_scatter_jj_gene <- df_scatter_jj %>%
  group_by(SYMBOL, Mean_expression_tissues, Intensity, HR) %>%
  summarise(
    Status = ifelse(any(Status == "New"), "New", "Known"),
    Rank = min(Rank),
    .groups = "drop"
  )

# Labeling genes
df_labels <- df_scatter_jj_gene %>%
  filter(Mean_expression_tissues < quantile(df_scatter_hr$Mean_expression_tissues, 0.75),
         Intensity > median(df_scatter_hr$Intensity)) %>%
  mutate(is_sb = Rank <= 0.5)

# Plot
p <- ggplot(df_scatter_jj_gene,
            aes(x = Mean_expression_tissues,
                y = Intensity,
                color = Status,
                text = paste0("Gene: ", SYMBOL,
                              "<br>Status: ", Status,
                              "<br>Mean expression: ", round(Mean_expression_tissues, 3),
                              "<br>Intensity: ", round(Intensity, 3),
                              "<br>HR: ", round(HR, 3),
                              "<br>Rank: ", round(Rank, 3)))) +
  geom_point(size = 2, alpha = 0.8) +
  geom_point(
    data = df_legend,
    aes(x = Mean_expression_tissues, y = Intensity, color = Status),
    size = 0, alpha = 0, inherit.aes = FALSE) +
  scale_color_manual(
  values = c("Known" = "#1f77b4",
             "New" = "#ff7f0e",
             "Strong binders" = "black",
             "Others" = "black"),
  labels = c("Known" = "Known peptides",
             "New"  = "Novel peptides",
             "Strong binders" = expression(bold("Strong binders (bold)")),
             "Others" = "Others"),
  breaks = c("Known", "New", "Strong binders", "Others")) +
  guides(color = guide_legend(
    override.aes = list(
      size  = c(2, 2, 0, 0),
      alpha = c(1, 1, 0, 0)))) +
  geom_hline(yintercept = median(df_scatter_hr$Intensity), linetype = "dashed", color = "black") +
  annotate("text", x = max(df_scatter_jj_gene$Mean_expression_tissues), y = median(df_scatter_hr$Intensity) - 0.2, label = "Median\nexpression", size = 2.5) +
  geom_vline(xintercept = quantile(df_scatter_hr$Mean_expression_tissues, 0.75), linetype = "dashed", color = "black") +
  annotate("text", x = quantile(df_scatter_hr$Mean_expression_tissues, 0.75) + 0.8, y = max(df_scatter_jj_gene$Intensity) + 0.2, label = "75th percentile", size = 2.5) +
  geom_text_repel(
    data = df_labels,
    aes(label = SYMBOL,
        fontface = ifelse(is_sb, "bold", "plain"),
        size     = 3),
    color         = "black",
    max.overlaps  = 20,
    box.padding   = 0.4,
    segment.color = "black",
    segment.size  = 0.3,
    min.segment.length = 0) +
  scale_size_identity() +
  scale_x_log10(breaks = c(0.001, 0.01, 0.1, 1, 10), labels = c("0.001", "0.01", "0.1", "1", "10")) +
  theme_minimal() +
  labs(x = "CTAs expression in normal tissues (Log10 TPM mean)",
       y = "CTAs expression in CHS (mean intensities)",
       color = "JJ012 HLA-A*02:01\npeptides")
p
#ggsave("../results/scatter_plot_jj.svg", plot = p,  width = 20, height = 15, units = "cm")
```
\newpage



### Rank scatter plot
```{r, fig.cap="Rank scatter plot (JJ012)"}
# Count number of peptides per rank
df_count <- df_scatter_jj %>%
  ungroup() %>%
  mutate(Status = as.character(Status)) %>%
  group_by(Rank) %>%
  summarise(
    nb_pep = n(),
    is_new = any(Status == "New"))

# Scatter plot
p <- ggplot(df_count, aes(x = nb_pep, y = Rank, color = is_new)) +
  geom_point(size = 2) +
  scale_color_manual(values = c("FALSE" = "grey30", "TRUE" = "#ff7f0e"), 
                     labels = c("FALSE" = "Known peptides", "TRUE" = "Novel peptides")) +
  geom_hline(yintercept = c(0.5, 2), linetype = "dashed", color = "black") +
  annotate("text", x = max(df_count$nb_pep), y = 0.5 * 0.8, label = "Strong binders", hjust = 1, color = "black") +
  annotate("text", x = max(df_count$nb_pep), y = 2 * 0.8, label = "Weak binders", hjust = 1, color = "black") +
  labs(x = "JJ012 HLA-A*02:01 peptides number", y = "%Rank", color = "Category") +
  scale_y_log10(breaks = c(0.01, 0.1, 1, 10, 100), labels = c("0.01", "0.1", "1", "10", "100")) +
  theme_minimal()
p
#ggsave("../results/scatter_jj_a02.svg", plot = p)
```
\newpage


# IV. Supplementary data 1
```{r}
# Select data
df_ch_supp <- df_scatter_ch[df_scatter_ch$Intensity > median(df_scatter_hr$Intensity) & df_scatter_ch$Mean_expression_tissues < quantile(df_scatter_hr$Mean_expression_tissues, 0.75), c("SYMBOL", "Intensity", "Mean_expression_tissues", "Cluster", "Seq", "Affinity", "Rank", "Status")]
df_jj_supp <- df_scatter_jj[df_scatter_jj$Intensity > median(df_scatter_hr$Intensity) & df_scatter_jj$Mean_expression_tissues < quantile(df_scatter_hr$Mean_expression_tissues, 0.75), c("SYMBOL", "Intensity", "Mean_expression_tissues", "Cluster", "Seq", "Affinity", "Rank", "Status")]

# Write in excel file
write_xlsx(list("JJ012_peptides_cadran" = df_jj_supp, "CH2879_peptides_cadran" = df_ch_supp), path = "../results/fig3_supp.xlsx")
```

